---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gbem

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/gbem)](https://CRAN.R-project.org/package=gbem)
[![R-CMD-check](https://github.com/stochaGBEM/gbem/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/stochaGBEM/gbem/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/stochaGBEM/gbem/branch/main/graph/badge.svg)](https://app.codecov.io/gh/stochaGBEM/gbem?branch=main)
<!-- badges: end -->

The goal of gbem is to calculate erosion volumes expected at a river's cross section, given flow magnitude and duration. 

**WARNING**: This package is under active development and is not intended to be used yet.

## Installation

As gbem is not on CRAN, you can install it from GitHub by running the following command in R:

``` r
remotes::install_github("stochaGBEM/gbem")
```

If this doesn't work, you may have to install the remotes package with the command `install.packages("remotes")`.

## Example

```{r example}
library(gbem)
```

### Hydrographs

Make a hydrograph from scratch:

```{r}
hg <- hydrograph(
  2 ~ 0, 3 ~ 1, 5 ~ 2, 3 ~ 5, 2 ~ 10
)
plot(hg)
```

Or, from a data frame; this time, also specify a time multiplier:

```{r}
data.frame(times = c(0, 1 / 3, 1), flow = c(100, 500, 200)) |> 
  as_hydrograph(times_from = times, flows_from = flow, unit = 24) |> 
  plot()
```

There are canned hydrographs for snowmelt-related and rainfall-related events, too:

```{r}
hydrograph_rain(peak = 200, baseflow = 50) |> 
  plot()
hydrograph_snow(peak = 200, baseflow = 50) |> 
  plot()
```

You can easily discretize the hydrograph if you'd like to do manual calculations:

```{r}
discretize_hydrograph(hg, 10)
```

### Cross Sections

Make a channel cross section:

```{r}
cs <- cross_section(width = 3, grad = 0.01, d50 = 0.1, d84 = 0.5, roughness = 0.01)
cs
```

### Gravel-bed river bank erosion model

Execute the algorithm for a constant flow by running `gbem()`. That returns a gbem object containing information about channel changes. Run `erode()` to execute the changes to get a new cross section:

```{r}
g1 <- gbem(100, t = 1, cross_section = cs)
erode(g1)
```

You can also iterate GBEM across a hydrograph. 

```{r}
g2 <- gbem2(hg, cross_section = cs, niter = 100)
erode(g2)
```

